---
title: "lab19"
author: "Nate Tran"
format: pdf
---

# Investigating pertussis cases by year

```{r}
library(datapasta)
library(ggplot2)
```

## Q1
Plot of pertussis cases by year shown below

Importing CDC pertussis data by year using datapasta

```{r}
cdc <- data.frame(
                                 Year = c(1922L,1923L,1924L,1925L,
                                          1926L,1927L,1928L,1929L,1930L,1931L,
                                          1932L,1933L,1934L,1935L,1936L,
                                          1937L,1938L,1939L,1940L,1941L,1942L,
                                          1943L,1944L,1945L,1946L,1947L,
                                          1948L,1949L,1950L,1951L,1952L,
                                          1953L,1954L,1955L,1956L,1957L,1958L,
                                          1959L,1960L,1961L,1962L,1963L,
                                          1964L,1965L,1966L,1967L,1968L,1969L,
                                          1970L,1971L,1972L,1973L,1974L,
                                          1975L,1976L,1977L,1978L,1979L,1980L,
                                          1981L,1982L,1983L,1984L,1985L,
                                          1986L,1987L,1988L,1989L,1990L,
                                          1991L,1992L,1993L,1994L,1995L,1996L,
                                          1997L,1998L,1999L,2000L,2001L,
                                          2002L,2003L,2004L,2005L,2006L,2007L,
                                          2008L,2009L,2010L,2011L,2012L,
                                          2013L,2014L,2015L,2016L,2017L,2018L,
                                          2019L),
         No..Reported.Pertussis.Cases = c(107473,164191,165418,152003,
                                          202210,181411,161799,197371,
                                          166914,172559,215343,179135,265269,
                                          180518,147237,214652,227319,103188,
                                          183866,222202,191383,191890,109873,
                                          133792,109860,156517,74715,69479,
                                          120718,68687,45030,37129,60886,
                                          62786,31732,28295,32148,40005,
                                          14809,11468,17749,17135,13005,6799,
                                          7717,9718,4810,3285,4249,3036,
                                          3287,1759,2402,1738,1010,2177,2063,
                                          1623,1730,1248,1895,2463,2276,
                                          3589,4195,2823,3450,4157,4570,
                                          2719,4083,6586,4617,5137,7796,6564,
                                          7405,7298,7867,7580,9771,11647,
                                          25827,25616,15632,10454,13278,
                                          16858,27550,18719,48277,28639,32971,
                                          20762,17972,18975,15609,18617)
       )
```

Constructing plot using ggplot2

```{r}
ggplot(cdc) +
  aes(Year, No..Reported.Pertussis.Cases) +
  geom_point() +
  geom_line() +
  labs(title="Pertussis Cases by Year", subtitle="1922-2019", x="Year", y="Number of Cases")
```

# Comparing Two Types of Pertussis Vaccines

## Q2
plot below

```{r}
ggplot(cdc) +
  aes(Year, No..Reported.Pertussis.Cases) +
  geom_point() +
  geom_line() +
  labs(title="Pertussis Cases by Year", subtitle="1922-2019", x="Year", y="Number of Cases") +
  geom_vline(xintercept=c(1946, 1996), linetype="dashed", col=c("blue", "red"))
```

## Q3
After introduction of the aP vaccine, pertussis cases seem to increase at a steady rate. The observed trend may be due to more sensitive testing methods compared to before the inception of PCR-based testing.

# Exploring CMI-PB Data

```{r}
library(jsonlite)
```

Loading in subject data form CMI-PB.

```{r}
subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 
```

## Q4
There are 47 aP and 49 wP infancy vaccinated subjects
```{r}
table(subject$infancy_vac)
```

## Q5
There are 66 female and 30 male subjects
```{r}
table(subject$biological_sex)
```

## Q6
Race and biological sex breakdown in table below
```{r}
table(subject$biological_sex, subject$race)
```

# Working with Dates

```{r}
library(lubridate)
library(dplyr)
```

## Q7
The average age of wP individuals is 36 years, and the average age of aP individuals is 26; these are significantly different
```{r}
subject$age <- time_length(today() - ymd(subject$year_of_birth), "years")

ap <- subject %>% filter(infancy_vac=="aP")
wp <- subject %>% filter(infancy_vac=="wP")

round(summary(ap$age))
round(summary(wp$age))
```

## Q8
The age of all individuals at time of boost is stored as the last column of subject object now
```{r}
subject$age_at_boost <- time_length(ymd(subject$date_of_boost) - ymd(subject$year_of_birth), "years")
```

## Q9
These two groups are significantly different, as shown by the histogram below.
```{r}
ggplot(subject) +
  aes(x=age, fill=as.factor(infancy_vac)) +
  geom_histogram(show.legend=F) +
  facet_wrap(vars(infancy_vac), nrow=2)
```

# Joining Multiple Tables

Reading and storing specimen and ab_titer tables

```{r}
specimen <- read_json("https://www.cmi-pb.org/api/specimen", simplifyVector = TRUE) 
titer <- read_json("https://www.cmi-pb.org/api/ab_titer", simplifyVector = TRUE) 
```

## Q10
Code for joining tables below
```{r}
meta <- left_join(specimen, subject)
abdata <- inner_join(titer, meta)
```

## Q11
6698 specimens for IgE, 1413 for IgG, 6141 for IgG1, IgG2, IgG3, IgG4
```{r}
table(abdata$isotype)
```

## Q12
There are signficantly fewer specimens collected on visit 8 than previous visits
```{r}
table(abdata$visit)
```

# Examining IgG1 Ab Titer Levels

Filtering for IgG1 isotype and excluding the small number of visit8 entries

```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
```

## Q13
Boxplots below
```{r}
ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot() +
  facet_wrap(vars(visit), nrow=2)
```

## Q14
FIM2 and FIM3 show significant changes over time. FHA and FRN show modest increases over time. 

Faceting boxplot by type of infancy vaccine

```{r}
ggplot(ig1) +
  aes(MFI, antigen, col=infancy_vac ) +
  geom_boxplot(show.legend = FALSE) + 
  facet_wrap(vars(infancy_vac, visit), nrow=2)
```

## Q15
Boxplots below
```{r}
filter(ig1, antigen=="FHA") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend =F) +
  facet_wrap(vars(visit)) +
  theme_bw() +
  labs(title="FHA antigen levels  per visit", subtitle="aP in red, wP in teal")

filter(ig1, antigen=="FIM2/3") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend =F) +
  facet_wrap(vars(visit)) +
  theme_bw() +
  labs(title="FIM2/3 antigen levels  per visit", subtitle="aP in red, wP in teal")
```

## Q16
FIM2/3 levels clearly rise over time and far exceed those of other antigens such as FHA. They peak at visit 5 and then decline, a trend seen in both aP and wP subjects.

## Q17
The FIM2/3 antigen time course shows that those vaccinated with aP vaccines have lower Ab titer response to FIM2/3 than those vaccinated with wP.

# Obtaining CMI-PB RNASeq

```{r}
url <- "https://www.cmi-pb.org/api/v2/rnaseq?versioned_ensembl_gene_id=eq.ENSG00000211896.7"

rna <- read_json(url, simplifyVector = TRUE)
```

Joining meta and rna data tables

```{r}
ssrna <- inner_join(rna,meta)
```

## Q18
Plot below
```{r}
ggplot(ssrna) +
  aes(visit, tpm, group=subject_id) +
  geom_point() +
  geom_line(alpha=0.2)
```

## Q19
The expression of IGHG1 reaches its maximum level at visit 4

## Q20
No, this pattern does not match the trend of FIM2/3 antibody titer data exactly, but this makes sense as the cells would be making more antibody that would facilitate the maximal response (highest Ab titer against antigens) seen by visit 5.

```{r}
ssrna %>%  
  filter(visit==4) %>% 
  ggplot() +
    aes(tpm, col=infancy_vac) + geom_density() + 
    geom_rug()
```

