---
title: "Lab11 Part 2"
author: "Nate Tran"
format: pdf
---

# Section 4: Population Scale Analysis

Reading file into environment and preliminarily exploring data.

```{r}
data <- read.table("pop_rs8067378.txt")
summary(data)
unique(data$geno)
```

## Q13
There are: 
  108 samples with genotype A/A, with a median expression value of 31.25.
  233 samples with genotype A/G, with a median expression value of 25.06.
  121 samples with genotype G/G, with a median expression value of 20.07.

```{r}
#extracting indices for each genotype sample group
idx_AA <- grep("A/A", data$geno)
idx_AG <- grep("A/G", data$geno)
idx_GG <- grep("G/G", data$geno)

#determining sample size for each genotype
length(idx_AA)
length(idx_AG)
length(idx_GG)

#calculating median expression levels for each genotype
median(data$exp[idx_AA])
median(data$exp[idx_AG])
median(data$exp[idx_GG])
```

## Q14
Based on the below boxplot, we can infer that each G substitution at the SNP position progressively decreases the expression of ORMDL3. The SNP and its different allele states definitely affects ORMDL3 expression.

```{r}
library(ggplot2)

ggplot(data) +
  aes(geno, exp) +
  geom_boxplot(alpha=0.4) +
  geom_jitter(alpha=0.4)
```

