---
title: "lab10"
author: "Nate Tran"
format: pdf
---
#Candy Dataset
```{r}
candy <- read.csv("candy-data.csv", row.names=1)
library(skimr)
skim(candy)
```

## Q1
There are 85 candy types in the dataset.
```{r}
dim(candy)
```


## Q2
There are 38 fruity candy types in the dataset.
```{r}
sum(candy$fruity)
```

## Q3
Favorite is Rolo and it has a winpercent of 65.7%.
```{r}
candy["Rolo",]$winpercent
```

## Q4
The winpercent of Kit Kat is 76.8%

```{r}
candy["Kit Kat",]$winpercent
```

## Q5
Thee winpercent of Tootsie Roll Snack Bars is 49.7%.
```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

## Q6
winpercent seems to be on a different scale to the majority of the other columns.
```{r}
skim(candy)
```

## Q7
A zero represents that the candy type is not chocolate and a one represents that it is chocolate.

## Q8
Plotted below
```{r}
hist(candy$winpercent)
```

## Q9
The distribution is not symmetrical.

## Q10
The center of the distribution is below 50%.

## Q11
Chocolate candy is higher ranked than fruity candy on average.
```{r}
win.choc <- candy$winpercent[as.logical(candy$chocolate)]
win.fruit <- candy$winpercent[as.logical(candy$fruity)]
mean(win.choc) > mean(win.fruit)
```

## Q12
Yes, this difference is statistically significant.
```{r}
t.test(win.choc, win.fruit)
```

# Overall Candy Rankings

## Q13
The five least liked candies are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters.
```{r}
ord.idx <- order(candy$winpercent)
head(candy[ord.idx,])
```

## Q14
The five most liked candies are Reese's pieces, Snickers, Kit Kats, Twix, and Reese's Miniatures.
```{r}
tail(candy[ord.idx,])
```

# Making Useful barplots

## Q15

```{r}
library(ggplot2)
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

# Coloring barplots

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "red"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) +
  ylab("Candy Type")
```

## Q17
The worst ranked chocolate candy is Sixlets

## Q18
The best ranked fruity candy is Starburst

# Pricepercent

## Q19
The highest ranked candy for the least amount of money is Reese's Miniatures.
```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

## Q20
The least popular of the most expensive candies is Nik L Nip.
```{r}
ord_idx_price <- order(candy$pricepercent, decreasing=T)
head(candy[ord_idx_price,11:12])
```

# Exploring Correlation

## Q22
Fruity and chocolate are anti-correlated

## Q23
Chocolate and winpercent are most positively correlated.
```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

# PCA
```{r}
pca <- prcomp(candy, scale=T)
summary(pca)

pc_plot <- ggplot(as.data.frame(pca$x)) +
  aes(PC1, PC2) +
  geom_point(size=candy$winpercent/10, col=my_cols) +
  geom_text_repel(label=rownames(candy), col=my_cols) +
  theme(legend.position="none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)

##ggplotly(pc_plot)
```

## Q24
Fruity, hard, and bar are captured by PC1.
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

